### 2025-07-02 README.Rmd
New ReadME

## Example

@@

```{r example, echo=TRUE, eval=FALSE }
# a constant TAC example
starttime <- (Sys.time())
options("show.signif.stars"=FALSE,
        "stringsAsFactors"=FALSE,
        "max.print"=50000,
        "width"=240)
# declare libraries ------------------------------------------------------------
suppressPackageStartupMessages({
  library(aMSE)       # from github.com/haddonm/aMSE MSE codebase
  library(EGHS)       # from github.com/haddonm/EGHS example Tas harvest strategy
  library(codeutils)  # from github.com/haddonm/codeutils a set of code utilities
  library(hplot)      # from github.com/haddonm/hplot a set of plotting routines
  library(makehtml)   # from github.com/haddonm/makehtml used to display results
  library(knitr)      # from CRAN for generating tables
})
dbdir <- getDBdir()   # utility to define the dropbox directory
# Obviously you should modify the rundir and datadir to suit your own setup
prefixdir <- paste0(dbdir,"A_CodeUse/aMSEUse/scenarios/")
postfixdir <- "M15h75"      # name your scenario, here natural mortality=0.15, steepness=0.75 
rundir <- paste0(prefixdir,postfixdir) # files and results directory
confirmdir(rundir)
controlfile <- "controlM15h75.csv"    # name of the control file
# equilibrium zone -------------------------------------------------------------
# You now need to ensure that there is, at least, a control??.csv, and a 
# saudata??.csv file in the data directory plus some other data .csv files
# depending on how conditioned you want the model to be. Templates for the
# correct format can be produced using ctrlfiletemplate(), datafiletemplate().
# 
# Of course, usually one would use data files, control.csv and a saudata.csv, which
# is listed as the datafile within the control.csv. These must be stored in 
# rundir. Read the Using_aMSE' section in the Documentation .docx file in the 
# documentation. To see examples of code to run the MSE for a real question.

# Define the arguments used in the harvest strategy
hsargs <- list(mult=0.1,
               wid = 4,
               targqnt = 0.55,
               maxtarg = c(150,150,150,150,150,150,150,150),
               pmwts = c(0.65,0.25,0.1),
               hcr = c(0.25,0.75,0.8,0.85,0.9,1,1.05,1.1,1.15,1.2),
               hcrm3 = c(0.25,0.75,0.8,0.85,0.9,1,1.1,1.15,1.2,1.25),
               startCE = 1992,
               endCE = 2020,
               metRunder = 2,
               metRover = 2,
               decrement=1,
               pmwtSwitch = 4,
               stablewts = c(0.4, 0.5, 0.1),
               hcrname="constantrefhcr",
               printmat=NULL)

@@

deleteyrs <- 0  # none removed in the western zone

# One would need to unhash this to get it to run (obviously)

#out <- do_MSE(rundir,controlfile,hsargs=hsargs,hcrfun=constantrefhcr,
#              sampleCE=tasCPUE,sampleFIS=tasFIS,sampleNaS=tasNaS,
#              getdata=tasdata,calcpopC=calcexpectpopC,makeouthcr=makeouthcr,
#              fleetdyn=NULL,scoreplot=plotfinalscores,
#              plotmultflags=plotmultandflags,interimout="",
#              varyrs=7,startyr=38,verbose=TRUE,ndiagprojs=4,cutcatchN=56,
#              matureL=c(40,170),wtatL=c(50,200),mincount=120,
#              includeNAS = FALSE,depensate=0,kobeRP=c(0.4,0.2,0.15),
#              nasInterval=1,minsizecomp=c(100,135),uplimH=0.375,incH=0.005,
#              deleteyrs=deleteyrs,selectyr=0)

#makeoutput(out,rundir,postfixdir,controlfile,hsfile="TasHS Package",
#           openfile=TRUE,verbose=FALSE)

# save(out,file=paste0(outdir,postfixdir,".RData"))

@@

After running the whole, even if you do not generate the results webpage, you could also try:

```{r }
#str(out$zoneDP,max.level=1)
#str1(out$hcrout)

To see the structure of the dynamic object generated by the projections.

>>>>

## Constant Reference Years Harvest Strategy

@@

```{r example, echo=TRUE, eval=FALSE }
# The following is copied from aMSEGuide Chapter 3
starttime <- (Sys.time())
options("show.signif.stars"=FALSE,
        "stringsAsFactors"=FALSE,
        "max.print"=50000,
        "width"=240)
# declare libraries ------------------------------------------------------------
suppressPackageStartupMessages({
  library(aMSE)       # from github.com/haddonm/aMSE MSE codebase
  library(EGHS)       # from github.com/haddonm/EGHS example Tas harvest strategy
  library(codeutils)  # from github.com/haddonm/codeutils a set of code utilities
  library(hplot)      # from github.com/haddonm/hplot a set of plotting routines
  library(makehtml)   # from github.com/haddonm/makehtml used to display results
  library(knitr)      # from CRAN for generating tables
})
# The next four code lines define my directory structure. 
# OBVIOUSLY, modify the rundir definition to suit your own setup!!!
dbdir <- getDBdir()   # utility to identify the dropbox directory
prefixdir <- pathtopath(dbdir,"A_codeR/aMSEGuide/runs/")
postfixdir <- "EG"     # a descriptive name for the rundir 
rundir <- pathtopath(prefixdir,postfixdir) # define rundir
startime <- Sys.time() # to document the time taken
verbose <- TRUE        # send messages to the console about the run progress
controlfile <- paste0("control",postfixdir,".csv") # match control file name
outdir <- "C:/aMSE_scenarios/EG/"   # storage on a non-cloud hard-drive
confirmdir(rundir,ask=FALSE)   # make rundir if it does not exist
confirmdir(outdir,ask=FALSE)   # to be interactive needs ask = TRUE
# You now need to ensure that there is, at least, a control??.csv, and a 
# saudata??.csv file in the data directory plus some other data .csv files
# Templates for the correct format can be produced using ctrlfiletemplate()
# and datafiletemplate().
# 
# Of course, usually one would use data files, control.csv and a saudata.csv, 
# which is listed as the datafile within the control.csv. These must be stored 
# in rundir. Read the Using_aMSE' section in aMSEGuide.
controlfile <- "controlEG.csv" # default example file name, change as needed
# Now make the controlfile. Read the help file using ?ctrlfiletemplate. This
# explains what the devrec argument is all about.

ctrlfiletemplate(indir=rundir,filename=controlfile,devrec=0)
# Within controlfile, the default data file name is 'saudatapostfixdir.csv'. If
# you want to call it something else (maybe aloysius.csv or machynlleth.csv?)
# then edit the seventh line of the controlfile that holds the definition
# of the name of the SAU data file, and change it in this code.

datafiletemplate(indir=rundir,filename="saudataEG.csv")
# The program also needs some length composition data. For this we are going 
# to use one of the inbuilt data-sets called 'lfs' (see its help). The default
# filen is already written into the controlfile. Change that as necessary.

data(lfs)
writecompdata(indir=rundir,lfs,filen="lf_WZ90-20.csv")

# Define the arguments used in the constantrefhcr harvest strategy
hsargs <- list(mult=0.1, #expansion factor for cpue range when calc the targqnt
               wid = 4, # number of years in the grad4 PM
               targqnt = 0.55, # quantile defining the cpue target
               maxtarg = c(150,150,150,150,150,150,150,150), # max cpue Target
               pmwts = c(0.65,0.25,0.1),  # relative weights of PMs
               hcr = c(0.25,0.75,0.8,0.85,0.9,1,1.05,1.1,1.15,1.2), # hcr mults
               hcrm3 = c(0.25,0.75,0.8,0.85,0.9,1,1.1,1.2,1.25,1.3),# mr 3
               startCE = 2000, # used in constant reference period HS
               endCE = 2019,   # used in constant reference period HS
               metRunder = 0,  # should the metarules be used. 0 = No
               metRover = 0,   # use metarules 0 = No
               decrement=1, # use fishery data up to the end of the time series
               pmwtSwitch = 0, # n years after reaching the targCE to replace
               stablewts = c(0.8, 0.15, 0.05), # pmwts with stablewts mr3
               hcrname="constantrefhcr",     # the name of the HCR used
               printmat=NULL) # something needed for some HS, not TAS

hsargs <- checkhsargs(hsargs) # lists chosen options before run, only for TAS 
# if you want to check your own set then include a suitable check function 
# within your jurisdiction's HS source file or package (see EGHS as example).

deleteyrs <- 0  # no years of data removed from the zone

# ?do_MSE provides more detailed descriptions of each function argument
# see EGHS package for help with the HS functions
# do_MSE function can take a few minutes to run so be patient
out <- do_MSE(rundir,controlfile, # already known
              hsargs=hsargs,    # defined as global object
              hcrfun=constantrefhcr,   # the main HS function
              sampleCE=tasCPUE, # processes cpue data
              sampleFIS=tasFIS, # processes FIS data (if any)
              sampleNaS=tasNaS, # processes Numbers-at-Size data
              getdata=tasdata,  # extracts the data from the zoneDP object
              calcpopC=calcexpectpopC, #distributes catches to populations
              makeouthcr=makeouthcr, # generates updateable HS output object
              fleetdyn=NULL,   # only used in SA and VIC so far
              scoreplot=plotfinalscores, # a function to plot HS scores
              plotmultflags=plotmultandflags, # function to plot multipliers
              interimout="", # save the results after projecting,
              varyrs=7, # years prior to projections for random recdevs
              startyr=48, # in plots of projections what year to start
              verbose=TRUE, # send progress reports to the console
              ndiagprojs=4, # individual trajectories in DiagProj plots
              cutcatchN = 56,  # cut size-at-catch array < class = 112mm
              matureL = c(70, 200), # length range for maturity plots
              wtatL = c(50, 200),   # length range for weight-at-length plots
              mincount = 120,   # minimum sample size in size-composition data
              includeNAS = FALSE, #include the numbers-at-size in saved output?
              depensate=0, #will depensation occur? see do_MSE help for details
              kobeRP=c(0.4,0.2,0.15), # ref points for a kobe-like phase plot
              nasInterval=5,  # year interval to use when plotting pred NaS 
              minsizecomp=c(100,135), #min size for pred sizes, min for catches
              uplimH=0.35,incH=0.005, #H range when estimating productivity 
              deleteyrs=0)    # all length comp years used

makeoutput(out,rundir,postfixdir,controlfile,
           hsfile="EGHS Package",openfile=TRUE,verbose=FALSE)

# unhash this to save the output to your outdir
# save(out,file=paste0(outdir,postfixdir,".RData"))

After running the whole, even if you do not generate the results webpage, to see the structure of the dynamic object generated by the projections you could try:

```{r, echo=TRUE, eval=FALSE }
str(out$zoneDP,max.level=1)
str1(out$hcrout)

### Constant Catch Harvest Strategy

Within EGHS there are two harvest control rules. The code above uses the constant reference years harvest strategy 'constantrefhcr'. There is also the 'consthr' which simply implements a constant unvarying aspirational catch in each SAU each year. To run that you should first copy the required data files into a new directory. As before, setup the R code to match whatever directory structure you intend to use:

```{r,  echo=TRUE, eval=FALSE}
library(codeutils)
library(aMSE)
dropdir <-getDBdir()   # once again FIX this to suit your own directory setup.
prefixdir <- pathtopath(dropdir,"/A_codeR/aMSETas/scenarios/")
filelist=c("controlEG.csv","lf_WZ90-20.csv","saudataEG.csv")
aMSE::copyto(prefixdir=prefixdir,fromdir="EG",todir="EGconst",
              filelist=filelist)
# this copies the files across and renames them appropriately.
```

The change in the harvest strategy is implemented in the hsargs rather than in the code, although it is necessary to let the do_MSE() funciton know of the change:

```{r,  echo=TRUE, eval=FALSE}
# amend the run code above with the following
postfixdir <- "EGconst"
verbose <- TRUE
rundir <- pathtopath(prefixdir,postfixdir)
controlfile <- paste0("control",postfixdir,".csv")
outdir <- "C:/aMSE_scenarios/EG/"
confirmdir(rundir,ask=FALSE)
confirmdir(outdir,ask=FALSE)

# current Tas HS
hsargs <- list(mult=0.025, # cpue range expansion factor when calc the targqnt
               wid = 4, # number of years in the grad4 PM
               targqnt = 0.55, # quantile defining the cpue target
               maxtarg = c(150,150,150,150,150,150,150,150), # max cpue Target
               acatch =  c(22,48,26,125,105,210,175,47),#ONLY used for consthcr
               pmwts = c(0.65,0.25,0.10),  # relative weights of PMs default
               hcr = c(0.25,0.75,0.8,0.85,0.9,1,1.05,1.1,1.15,1.2), 
               hcrm3 = c(0.25,0.75,0.8,0.85,0.9,1,1.1,1.2,1.25,1.3),
               #hcr = c(1,1,1,1,1,1,1,1,1,1,1), # multipliers for constant catch
               startCE = 2000, # used in constant reference period HS
               endCE = 2019,   # used in constant reference period HS
               #refperiodCE = 2000:2019, # use a vector of years instead
               metRunder = 2,  # MR2 if CE below targ multTAC = 1
               metRover = 2,   # MR1 CE>targ for 2+ yrs before multTAC increases
               decrement=1, # use fishery data up to the end of the time series
               pmwtSwitch = 4, # number of years after reaching the targCE to
               stablewts = c(0.65, 0.25, 0.1), # replace pmwts with stablewts
               hcrname="consthcr", # name of HCR: constantrefhcr / consthcr
               printmat=NULL)

@@

out <- do_MSE(rundir,controlfile,hsargs=hsargs,hcrfun=consthcr,  # constantrefhcr,     #
              sampleCE=tasCPUE,sampleFIS=tasFIS,sampleNaS=tasNaS,
              getdata=constdata,calcpopC=calcexpectpopC,makeouthcr=makeoutconst,
              fleetdyn=NULL,scoreplot=plotfinalscores,
              plotmultflags=plotmultandflags,interimout="",
              varyrs=7,startyr=38,verbose=TRUE,ndiagprojs=4,cutcatchN=56,
              matureL=c(40,170),wtatL=c(50,200),mincount=120,
              includeNAS = FALSE,depensate=0,kobeRP=c(0.4,0.2,0.15),
              nasInterval=1,minsizecomp=c(100,135),uplimH=0.35,incH=0.005,
              deleteyrs=0)

makeoutput(out,rundir,postfixdir,controlfile,hsfile="EGHS Package",
           openfile=TRUE,verbose=FALSE)

@@

### Final Words
